---
# Basic template for FMX Prac:

title: "Question 2: Currency Hedging Analysis"
subtitle: "Question 2: Currency Hedging Analysis"

documentclass: "elsarticle"

Thesis_FP: FALSE

# --------- Front Page

AddTitle: TRUE
Author1: "Jan-Hendrik Pretorius" 
Ref1: "Stellenbosch University" 
Email1: "20713479\\@sun.ac.za" 

# ----- Manage headers and footers:
BottomRFooter: "\\footnotesize Page \\thepage"
addtoprule: TRUE
addfootrule: TRUE               

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE

# --------- Line numbers
linenumbers: FALSE

# ---------- References settings:
bibliography: Tex/ref.bib       
csl: Tex/harvard-stellenbosch-university.csl

# ---------- General:
RemovePreprintSubmittedTo: TRUE  
Journal: "Financial Econometrics 871"
toc: FALSE                       
numbersections: TRUE             
fontsize: 11pt                  
linestretch: 1.2                
link-citations: TRUE            

### Adding additional latex packages:

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This report examines the impact of currency hedging on a 60/40 Equity/Bond portfolio. Comparing hedged and unhedged strategies, we analyze their effects on portfolio volatility and performance, providing insights for effective currency risk management in portfolio construction.
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

    #   Housekeeping
rm(list = ls())
options(scipen = 999)
path <- "/Users/janhendrikpretorius/Library/CloudStorage/OneDrive-StellenboschUniversity/01-Masters-2023/02 Financial Econometrics/20713479-FMX-Prac/"

    #   Get external scripts
source(paste0(path, "UTILITIES/libraries.R")) #   Load libraries
source(paste0(path, "UTILITIES/aesthetics.R")) #  Get plot themes
source(paste0(path, "UTILITIES/calc-portfolio-ret.R")) #  Function to calc portfolio returns
source(paste0(path, "UTILITIES/missing-values.R")) #. Function to impute missing values

    #   Load Data
file <- "Cncy_Hedge_Assets.rds"
Indexes <- read_rds(paste0(path, "DATA/", file))

file <- "Monthly_zar.rds"
ZAR <- read_rds(paste0(path, "DATA/", file))
```

<!-- ############################## -->

<!-- # Start of Question: -->

<!-- ############################## -->

```{r portfolio-returns}

RebMonths <- c(3, 6, 9, 12) # Rebalancing periods

# Convert the date format to Year-Month for both dataframes
Indexes <- Indexes %>% 
  mutate(date = floor_date(as.Date(date), "month"))

ZAR <- ZAR %>% 
  mutate(date = floor_date(as.Date(date), "month"))

# Note on calculatePortfolioReturns():
    # Need a wide format dataframe with a date column and returns only!
    # It is a very situational specific function and not very general

# Calculate returns of unhedged portfolio
Unhedged_portf <- calculatePortfolioReturns(Indexes, RebMonths) %>% 
    mutate(Structure = "Unhedged")

# Calculate returns of hedged portfolio

# Align dates and compute growth rate of ZAR to USD exchange rate
ZAR <- ZAR %>% mutate(ExchangeRateGrowth = (value / lag(value)) - 1)

# Adjust returns in Indexes by exchange rate appreciation/depreciation
Indexes_adjusted <- Indexes %>%
  mutate(date = as.Date(date)) %>%
  left_join(ZAR %>% select(date, ExchangeRateGrowth), by = "date") %>%
  mutate_at(vars(J433, ALBI), ~ . * (1 + ExchangeRateGrowth)) %>% # Inflate Rand-denominated indexes
  mutate_at(vars(MSCI_ACWI, Bbg_Agg), ~ . / (1 + ExchangeRateGrowth)) %>% # Deflate Dollar-denominated indexes
  select(-c(ExchangeRateGrowth))

Hedged_portf <- calculatePortfolioReturns(Indexes_adjusted, RebMonths) %>% 
    mutate(Structure = "Hedged")

# Merge the unhedged and hedged portfolio returns
combined_portfolio <- bind_rows(Unhedged_portf, Hedged_portf)

# Calculate cumulative and rolling returns
combined_portfolio <- combined_portfolio %>%
  group_by(Structure) %>%
  arrange(date) %>%
  mutate(CumulativeReturn = cumprod(1 + PortfolioReturn)) %>%
  mutate(RollRets = RcppRoll::roll_prod(1 + PortfolioReturn, 36, fill = NA, 
  align = "right")^(12/36) - 1) %>% 
  group_by(date) %>% filter(any(!is.na(RollRets))) %>% 
  ungroup()

```

# Hedged vs Unhedged Growth

```{r cumrets}

# Plot the cumulative returns
combined_portfolio %>% 
ggplot(aes(x = date, y = CumulativeReturn, color = Structure)) +
  geom_line() +
  th +
  labs(title = "Cumulative Portfolio Returns: Hedged vs Unhedged",
       x = "Date",
       y = "Cumulative Return",
       color = "Structure") +
  scale_color_manual(values = palette)

```


The graph above illustrates the cumulative returns of hedged versus unhedged portfolios from 2005 onwards. Both strategies show growth over time, with the hedged portfolio occasionally outperforming the unhedged one. There is some divergence, particularly during periods of market volatility, but the long-term trajectory of both portfolios appears similar, indicating that hedging may not significantly impact cumulative returns over an extended period.

The 3-year rolling returns chart below for hedged and unhedged portfolios highlights fluctuations over time. Despite some periods where the hedged portfolio shows slightly lower volatility, overall, both portfolios follow a similar pattern. This suggests that hedging may not provide a clear advantage in smoothing returns over a three-year rolling period.

```{r rollrets}

# Plot the rolling returns
combined_portfolio %>% 
ggplot(aes(x = date, y = RollRets, color = Structure)) +
  geom_line() +
  th +
  labs(title = "3-Year Rolling Returns: Hedged vs Unhedged",
       x = "Date",
       y = "Rolling 3 year Returns (Ann.)",
       color = "Structure") +
  scale_color_manual(values = palette)

```


The figure below was replicated from a recent study around currency hedging. It shows how global allocation and the rand have a strong negative relationship since 2004. This means that when one goes up, the other goes down, and vice versa. This happens because both are influenced by the same global factors. For example, when there's positive risk sentiment, both global allocation and the rand tend to do well, and when sentiment is negative, they both suffer.

Also, notice that the rand tends to drop more significantly in value in a month than it appreciates. So, if you make a wrong hedge (on the right side of the plot), it can be much costlier than the benefit of getting it right (on the left side).

```{r scatterplot, fig.width=7, fig.height=4}

# Convert returns to USD-returns
Indexes_USD <- Indexes %>%
  mutate(date = as.Date(date)) %>%
  left_join(ZAR %>% select(date, ExchangeRateGrowth), by = "date") %>%
  mutate_at(vars(J433, ALBI), ~ . / (1 + ExchangeRateGrowth)) %>% 
  select(-c(ExchangeRateGrowth))

USD_portf <- calculatePortfolioReturns(Indexes_adjusted, RebMonths) %>% 
    mutate(Hedged = "Yes")

scatter_data <- USD_portf %>%
  left_join(ZAR %>% select(date, ExchangeRateGrowth), by = "date") %>%
  rename(USD_ZAR_Returns = ExchangeRateGrowth,
         Portfolio_Returns_USD = PortfolioReturn) %>%
  select(date, USD_ZAR_Returns, Portfolio_Returns_USD)

# Create the scatter plot
p <- scatter_data %>% 
        ggplot(aes(x = USD_ZAR_Returns, y = Portfolio_Returns_USD)) +
            geom_point(color = "#D98515", alpha=0.6) +
            geom_smooth(method = "lm", color = "darkgrey", se=FALSE) +
            scale_x_continuous(labels = scales::percent_format(scale = 100), limits = c(-0.15, 0.15)) +
            scale_y_continuous(labels = scales::percent_format(scale = 100), limits = c(-0.15, 0.15)) +
            geom_hline(yintercept = 0, linetype = "dashed", color = "darkgrey") +
            geom_vline(xintercept = 0, linetype = "dashed", color = "darkgrey") +
            th +
            labs(x = "USD-ZAR Returns", 
                 y = "Portfolio Returns (USD)", 
                 title = "Scatter Plot with Marginal Distributions",
                 caption = "Calculations from 28 February 2002 - 31 August 2023")

# Add marginal histograms
p <- ggExtra::ggMarginal(p, type = "density", fill = "#1E3364", alpha = 0.6)

print(p)
```


# Volatility

When comparing the hedged and unhedged portfolios, the hedged portfolio tends to exhibit slightly higher downside risk, as indicated by the higher values in semi-deviation, downside deviation, maximum drawdown, VaR, ES, and modified VaR. However, the differences between the two portfolios are relatively small, suggesting that the choice between hedging or not may depend on other factors such as investment goals and risk tolerance. This information is presented in the table below.

```{r}

# Separate hedged and unhedged data
df_hedged <- combined_portfolio %>% filter(Structure == "Hedged") %>% select(date, PortfolioReturn)
df_unhedged <- combined_portfolio %>% filter(Structure == "Unhedged") %>% select(date, PortfolioReturn)

# Rename columns to match the example structure
df_hedged <- df_hedged %>% rename(Hedged = PortfolioReturn)
df_unhedged <- df_unhedged %>% rename(Unhedged = PortfolioReturn)

# Calculate Downside Risk
tabdownside <- table.DownsideRisk(left_join(df_unhedged, df_hedged, by = "date") %>% tbl_xts(.), 
                                  ci = 0.95, Rf = 0, MAR = 0)

# Select specific rows
tabdownside <- tabdownside[c(1,5,7,8:11),]

# Format and display the table
tabdownside %>% 
  data.frame() %>% 
  tibble::rownames_to_column() %>% 
  gt() %>% 
  tab_header(title = glue("Downside Risk Estimates")) %>% 
  fmt_percent(columns = 2:3, decimals = 2)



```


The graphical representations presented in the following figures (which include the 36-month rolling volatility, volatility histogram, and violin plot) provide compelling evidence that hedging strategies may not deliver the anticipated reduction in volatility. In fact, the data suggests a counterintuitive trend—over extended timeframes, volatility tends to exhibit an upward trajectory. This unexpected pattern challenges the fundamental goal of hedging, which is often implemented to mitigate risk and stabilize portfolio performance. This intriguing finding invites further exploration and analysis into the dynamics of hedging and its long-term impact on portfolio volatility.

```{r rolling-ret}

# Calculate 30-day rolling volatility
combined_portfolio <- combined_portfolio %>%
    group_by(Structure) %>%
    arrange(date, .by_group = TRUE) %>%
    mutate(RollingVolatility = rollapply(PortfolioReturn, width = 36, FUN = sd, na.rm = TRUE, fill = NA, align = "right"))

# Plotting
ggplot(combined_portfolio, aes(x = date, y = RollingVolatility, color = Structure)) +
    geom_line() +
    scale_color_manual(values = palette) +
    labs(title = "36-month Rolling Volatility",
         x = "Date",
         y = "Volatility (Standard Deviation)",
         color = "Structure") +
    th


```

```{r histogram}

volatility_data <- combined_portfolio %>%
  filter(!is.na(RollingVolatility))

# Create a histogram of the rolling volatility
ggplot(volatility_data, aes(x = RollingVolatility, fill = Structure)) +
  geom_histogram(position = "identity", alpha = 0.6, bins = 30, colour = "black", size = 0.3) +
  labs(title = "Histogram of 36-Month Rolling Volatility",
       x = "Rolling Volatility (Standard Deviation)",
       y = "Frequency") +
  scale_fill_manual(values = palette) +
  th

```


```{r}

# Create a box plot of the rolling volatility
ggplot(volatility_data, aes(x = Structure, y = RollingVolatility, fill = Structure, color = Structure)) +
  geom_violin(alpha = 0.6) +
  geom_jitter() +
  labs(title = "Violin Plot of 36-Month Rolling Volatility",
       x = "Portfolio Structure",
       y = "Rolling Volatility (Standard Deviation)") +
  scale_fill_manual(values = palette) +
  scale_color_manual(values = palette) +
  th

```

