---
# Basic template for FMX Prac:

title: "Question 5: Volatility and GARCH estimates"
subtitle: "Question 5: Volatility and GARCH estimates"

documentclass: "elsarticle"

Thesis_FP: FALSE

# --------- Front Page

AddTitle: TRUE
Author1: "Jan-Hendrik Pretorius" 
Ref1: "Stellenbosch University" 
Email1: "20713479\\@sun.ac.za" 

# ----- Manage headers and footers:
BottomRFooter: "\\footnotesize Page \\thepage"
addtoprule: TRUE
addfootrule: TRUE               

# --------- page margins:
margin: 2.3 
bottom: 2 
top: 2.5 
HardSet_layout: TRUE

# --------- Line numbers
linenumbers: FALSE

# ---------- References settings:
bibliography: Tex/ref.bib       
csl: Tex/harvard-stellenbosch-university.csl

# ---------- General:
RemovePreprintSubmittedTo: TRUE  
Journal: "Financial Econometrics 871"
toc: FALSE                       
numbersections: TRUE             
fontsize: 11pt                  
linestretch: 1.2                
link-citations: TRUE            

### Adding additional latex packages:

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 
    fig_height: 3.5
abstract: |
  This report analyzes currency market volatility and correlations, emphasizing the South African Rand (ZAR) and its major trading partners. Key insights include ZAR's volatility patterns, time-varying volatilities, and correlations with G10 currencies, offering valuable insights for risk management. I study the log returns of the ZAR for the period after the Global Financial Crisis, where ZAR saw increased volatility.
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

    #   Housekeeping
rm(list = ls())
options(scipen = 999)
path <- "/Users/janhendrikpretorius/Library/CloudStorage/OneDrive-StellenboschUniversity/01-Masters-2023/02 Financial Econometrics/20713479-FMX-Prac/"

    #   Get external scripts
source(paste0(path, "UTILITIES/libraries.R")) #   Load libraries
source(paste0(path, "UTILITIES/aesthetics.R")) #  Get plot themes
source(paste0(path, "UTILITIES/missing-values.R")) #  Get imputing function
source(paste0(path, "UTILITIES/capping.R")) #  Capping functions
source(paste0(path, "UTILITIES/renaming-ddc.R")) #  renamingddc function

    #   Load Data
file <- "currencies.rds"
cncy <- read_rds(paste0(path, "DATA/", file))

file <- "cncy_Carry.rds"
cncy_Carry <- read_rds(paste0(path, "DATA/", file))

file <- "cncy_value.rds"
cncy_value <- read_rds(paste0(path, "DATA/", file))

file <- "cncyIV.rds"
cncyIV <- read_rds(paste0(path, "DATA/", file))

file <- "bbdxy.rds"
bbdxy <- read_rds(paste0(path, "DATA/", file))

```

<!-- ############################## -->

<!-- # Start of Question: -->

<!-- ############################## -->

```{r data-operations}

# Prepare the data for volatility comparison

cncy <- cncy %>%
  group_by(Name) %>%
  arrange(date) %>% 
  mutate(LogReturns = log(Price) - log(lag(Price))) %>%
  mutate(ScaleRet = (LogReturns - mean(LogReturns, na.rm = T))) %>% 
  filter(!is.na(LogReturns)) %>%
  ungroup() %>% 
  filter(date > as.Date("2008-12-31")) # Filter dates to reflect currency movements after GFC

# Calculate the standard deviation as a measure of volatility
cncy_volatility <- cncy %>%
  group_by(Name) %>%
  summarize(AvgVolatility = sd(LogReturns, na.rm = TRUE))

# List of currencies to compare with ZAR
currencies_to_compare <- c("SouthAfrica_Cncy", "EU_Cncy_Inv", "UK_Cncy_Inv", "China_Cncy", "Japan_Cncy", "India_Cncy")

# Filter and calculate the standard deviation for each currency
selected_currencies_volatility <- cncy %>%
  filter(Name %in% currencies_to_compare) %>%
  group_by(Name) %>%
  summarize(AvgVolatility = sd(LogReturns, na.rm = TRUE)) %>%
  ungroup()

# Define the rolling window size
rolling_window <- 30  # e.g., 30-day rolling window

# Calculate rolling volatility for each currency
cncy_rolling_volatility <- cncy %>%
  filter(Name %in% currencies_to_compare) %>%
  group_by(Name) %>%
  mutate(RollingVolatility = rollapply(LogReturns, width = rolling_window, FUN = sd, na.rm = TRUE, fill = NA, align = "right")) %>%
  ungroup()

```



```{r barplot}
palette2 <- c("#1F6F70","#1F6F70", "#1F6F70","#1F6F70","#C93D44","#1F6F70")

# Visualize average volatility
ggplot(selected_currencies_volatility, aes(x = reorder(Name, AvgVolatility), y = AvgVolatility, fill = Name)) +
  geom_bar(stat = "identity", alpha = 0.6) +
  labs(title = "Average Volatility Comparison of ZAR with\nMajor Trading Partners", 
       x = "Currency", 
       y = "Average Volatility (Standard Deviation)") +
  coord_flip() +
  scale_fill_manual(values = palette2) +
  th +
  theme(legend.position = "none")


```

Compared to major trading partners, the Rand on average has seen much higher volatility over the past decade after the Global Financial Crisis (GFC) of 2008, exhibiting average volatility (standard deviation) of 0.01 - almost double the volatility faced by trading partner currencies.

```{r}
# Plot the rolling volatility
ggplot(cncy_rolling_volatility, aes(x = date, y = RollingVolatility, color = Name)) +
  geom_line(size = 0.5, alpha = 0.6) +
  labs(title = paste0(rolling_window, "-Day Rolling Volatility of ZAR and Major Trading\nPartners' Currencies"), 
       x = "Date", 
       y = "Rolling Volatility (Standard Deviation)",
       subtitle = "ZAR Rolling Volatility Highlighted in Red") +
  th +
  scale_color_manual(values = palette2) +
  theme(legend.position = "none")


```

The ZAR consistently exhibited higher 30-day rolling volatility than trading partners over the period under study. In the figure above, the ZAR is highlighted in red.

```{r fit-garch}

library(rugarch) # I define rugarch here because it messes with some of my other code if I define it in the libraries utility code
library(rmgarch)

# currencies_to_compare contains the list of currencies including ZAR
# Function to fit GARCH(1,1) model and return conditional volatility
fit_garch_model <- function(data, currency) {
  returns <- na.omit(data %>% filter(Name == currency) %>% select(LogReturns))
  spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
                     mean.model = list(armaOrder = c(0,0), include.mean = FALSE),
                     distribution.model = "std")
  fit <- ugarchfit(spec, returns$LogReturns)
  return(sigma(fit))
}

# Apply the function to each currency
volatility_list <- lapply(currencies_to_compare, function(currency) {
  data <- fit_garch_model(cncy, currency)
  data <- fortify(data)
  data$Currency <- currency
  return(data)
})

# Combine the volatility data into a single data frame
volatility_data <- bind_rows(volatility_list)

# Rename the 'data' column to 'Volatility' for clarity
volatility_data <- volatility_data %>%
  rename(Volatility = data)


```


Studying currency volatility involved fitting a GARCH(1,1) model to estimate time-varying volatilities of multiple currencies, including the South African Rand (ZAR). The model structure consisted of a GARCH(1,1) framework. This model is designed to capture volatility clustering and shocks in financial time series data. It works by modeling the conditional variance of returns as a function of past squared returns and past conditional variances. In simpler terms, it estimates how volatile each currency's returns are at any given time, taking into account its own historical volatility and squared returns, which helps identify periods of high or low volatility. The model was applied individually to each currency's returns, and the resulting conditional volatilities were used for comparative analysis.


```{r plot-garch}

# Plot the conditional volatilities
ggplot(volatility_data, aes(x = Index, y = Volatility, color = Currency)) +
  geom_line(size = 0.5, alpha = 0.6) +
  labs(title = "Time-Varying Volatility of Currencies (GARCH Model)", 
       x = "Date", 
       y = "Conditional Volatility") +
  th +
  scale_color_manual(values = palette2) +
  theme(legend.position = "none")

```

The graph above indicates the South African Rand (ZAR) experiences periods of higher volatility relative to its trading partners. These peaks suggest moments of economic stress or significant market events.

A GO-GARCH model methodology was applied in the analysis and serves to investigate the dynamic relationships between the South African Rand (ZAR) and the currencies of the G10 countries. The GO-GARCH, or Generalized Orthogonal GARCH, is a sophisticated variant of the standard GARCH model which allows for the assessment of time-varying correlations among multiple time series. In this context, the model captures how the volatility of the ZAR co-moves with the volatility of the G10 currency basket over time.

```{r go-garch}

G10_returns <- bbdxy %>% 
    mutate(G10 = log(Price)-log(lag(Price))) %>% 
    filter(date > dplyr::first(date)) %>% 
    select(date, G10)

# cncy is == to cncy_rts from the example code
ZAR <- cncy %>% 
  filter(Name == "SouthAfrica_Cncy") %>% 
  rename("ZAR" = "LogReturns") %>% 
  select(date, ZAR) 

# Merge ZAR and G10 returns
combined_returns <- left_join(G10_returns, ZAR, by = "date") %>% 
    filter(date > as.Date("2008-12-31")) %>% 
    tbl_xts()

# Set up the univariate GARCH specifications
uspec <- ugarchspec(variance.model = list(model = "gjrGARCH", garchOrder = c(1, 1)),
                    mean.model = list(armaOrder = c(1, 0), include.mean = TRUE),
                    distribution.model = "sstd")

# Replicate the specification for each series in combined_returns
multi_univ_garch_spec <- multispec(replicate(ncol(combined_returns), uspec))

# GO-GARCH model specification
spec.go <- gogarchspec(multi_univ_garch_spec, 
                       distribution.model = 'mvnorm', # or 'manig'
                       ica = 'fastica') # Using fastICA

# Fit the GO-GARCH model
fit.gogarch <- gogarchfit(spec.go, data = combined_returns, solver = 'hybrid')

# Extract time-varying correlations
gog.time.var.cor <- rcor(fit.gogarch)
gog.time.var.cor <- aperm(gog.time.var.cor, c(3,2,1))
dim(gog.time.var.cor) <- c(nrow(gog.time.var.cor), ncol(gog.time.var.cor)^2)
gog.time.var.cor <- renamingdcc(ReturnSeries = combined_returns, DCC.TV.Cor = gog.time.var.cor)


```



```{r plot-gog}
palette3 <- c("grey", "grey", "#C93D44", "grey")

# Plot the time-varying correlations
ggplot(gog.time.var.cor, aes(x = date, y = Rho, color = Pairs)) +
  geom_line() +
  labs(title = "Time-Varying Correlations: ZAR with G10 Currencies",
       x = "", 
       y = "Correlation") +
  th +
  scale_y_continuous(limits = c(NA, 0.75)) + # I'm gonna be honest here: I don't know how to filter for specific observations in gog.time.var.cor, so I just adjust the axis scale 
  scale_color_manual(values = palette3) +
  theme(legend.position = "none")

```

The plot above illustrates the time-varying correlations between the South African Rand (ZAR) and G10 currencies from 2010 to 2020. The correlations fluctuate over time, displaying a range predominantly between 0.55 and 0.75. Notably, there are peaks where the correlation approaches 0.75, suggesting periods where the ZAR moved more in tandem with the G10 currencies, possibly reflecting global economic events impacting markets uniformly.

Conversely, the valleys indicate times when the ZAR had lower correlations with the G10 currencies, which could imply periods where local factors or idiosyncratic events had a greater impact on the ZAR's movement, making it less aligned with the G10 currencies' trends.

```{r heat-map}

# Preparing the data
heatmap_data <- gog.time.var.cor %>%
  filter(grepl("ZAR_", Pairs), !grepl("_ZAR", Pairs)) %>%
  gather(key = "Metric", value = "Value", -date, -Pairs)

# Creating the heatmap
ggplot(heatmap_data, aes(x = date, y = Pairs, fill = Value)) +
  geom_tile() +
  scale_fill_gradient2(low = "#1F6F70", high = "#C93D44", mid = "white", midpoint = 0.5,
                       limit = c(-1, 1), space = "Lab", name="Correlation") +
  labs(title = "Heatmap of Time-Varying Correlations: ZAR with\nG10 Currencies",
       x = "", 
       y = "Currency Pair") +
  th +
  theme(legend.position = "bottom")


```

The heatmap depicts the time-varying correlations between the South African Rand (ZAR) and G10 currencies from 2010 to around 2020. The color gradient represents the strength of the correlation, where shades closer to white denote a correlation near zero (indicating no relationship), and darker shades, either red or blue, indicate stronger positive or negative correlations respectively.

In this heatmap, however, the color gradient seems to be on a scale from light to dark pink, suggesting that the correlations are predominantly positive throughout the period. The intensity of the pink shade varies over time, implying fluctuations in the degree of correlation. The consistent presence of color, without reverting to white, indicates that the ZAR generally maintains some level of positive correlation with the G10 currencies during the given timeframe.

There aren’t any dark red or blue patches, which would suggest extremely high or low correlations, and the lack of variability in color intensity indicates that while correlations fluctuate, they do so within a relatively narrow range. This pattern might suggest that while specific events or periods might affect the degree of co-movement between the ZAR and G10 currencies, the overall relationship remains relatively stable and positively correlated.
